Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-16 paper/us_election_2020.qmd
update appendix

## Data Result

@@

modelsummary(
  list(
    "Support Biden" = political_preferences1
  ),
  statistic = "mad"
  )
```
Our results are summarized in @tbl-modelresults1. Our results generally matches our expectation. To avoid multicollinearity, the model excluded three variables from each category: race Asian, region Midwest, and employed full-time.The intercept represents the estimated log-odds of supporting Biden when all other predictors are held constant at their reference levels. In this case, the estimated log-odds of supporting Biden for individuals who are Asian, live in the Midwest, and are employed full-time is 0.806.

@@

posterior_vs_prior(political_preferences1, size = 0.3) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

@fig-trace is a trace plot. It shows if there is existence of signs that the our model runs into issues. We observe lines that are horizontal and oscillating, and have overlaps between the chains. This suggests that there is nothing strange in this trace plot.

@fig-rhat is a Rhat plot. It compares the variability within each chain to the variability between chains in MCMC. Ideally, we want Rhat to be close to 1, which indicates that the chains have converged to the same distribution. We can observe that our Rhat plot are all close to 1, and no more than 1.05. This is a positive sign because it suggests that the MCMC algorithm has reached convergence for our model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace
#| fig-cap: "Trace plot"

 plot(political_preferences1, "trace")
```

@@

## Credibility Interval {#sec-credibility-interval}
90% Credibility intervals graph are shown in @fig-modelresults1 and @fig-modelresults2.
@fig-modelresults1 shows range of coefficient estimates of our model within the 90% probability. Due to the fact that the credibility interval for race "Other" is particularly large, we cannot observe the trend of the 90% credibility intervals of other variables. Therefore we created figure @fig-modelresults2 with the x axis limited from -10 to 10.

Combing @fig-modelresults1 and @fig-modelresults2, we observe statistical significance for the coefficient estimates for students, the Northeast region, individuals identifying with two or more races, individuals identifying with "Other" racial category, Native Americans, Black individuals, and the intercept, Asians in the Midwest region who are employed full-time. The estimates are significant if the intervals do not cross 0. The value for the estimates are in log-odds, indicating that if the coefficient is positive, the individual supports Biden, if negative, the individual supports Trump.

>>>>

## Data Results

@@

modelsummary(
  list(
    "Support Biden" = political_preferences1
  ),
  statistic = "mad"
  )
```
Our results are summarized in @tbl-modelresults1. Our results generally matches our expectation. To avoid multicollinearity, the model excluded three variables from each category: race Asian, region Midwest, and employed full-time. The intercept represents the estimated log-odds of supporting Biden when all other predictors are held constant at their reference levels. In this case, the estimated log-odds of supporting Biden for individuals who are Asian, live in the Midwest, and are employed full-time is 0.806.

@@

posterior_vs_prior(political_preferences1, size = 0.3) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  theme(axis.text.x=element_text(hjust=1, vjust=0.5),
        text=element_text(size=8)) +
  coord_flip()
```

@@

@fig-trace1, @fig-trace2, @fig-trace3 are the trace plot of the model. It shows if there is existence of signs that the our model runs into issues. We observe lines that are horizontal and oscillating, and have overlaps between the chains. This suggests that there is nothing strange in this trace plot.

@fig-rhat is the Rhat plot of the model. It compares the variability within each chain to the variability between chains in MCMC. Ideally, we want Rhat to be close to 1, which indicates that the chains have converged to the same distribution. We can observe that our Rhat plot are all close to 1, and no more than 1.05. This is a positive sign because it suggests that the MCMC algorithm has reached convergence for our model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace1
#| fig-cap: "Trace plot of intercept and race"
#| fig-subcap: ["Trace plot of Intercept", "Trace plot of race White", "Trace plot of race Black", "Trace plot of race Hispanic", "Trace plot of race Middle Eastern", "Trace plot of race Native American", "Trace plot of race Two or more races", "Trace plot of race Other"]
#| layout-ncol: 2

#plot(political_preferences1, "trace")
plot(political_preferences1, "trace", "(Intercept)")

plot(political_preferences1, "trace", "raceWhite")
plot(political_preferences1, "trace", "raceBlack")
plot(political_preferences1, "trace", "raceHispanic")
plot(political_preferences1, "trace", "raceMiddle Eastern")
plot(political_preferences1, "trace", "raceNative American")
plot(political_preferences1, "trace", "raceTwo or more races")
plot(political_preferences1, "trace", "raceOther")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace2
#| fig-cap: "Trace plot of region"
#| fig-subcap: ["Trace plot of region Northeast", "Trace plot of region South", "Trace plot of region West"]
#| layout-ncol: 2

plot(political_preferences1, "trace", "regionNortheast")
plot(political_preferences1, "trace", "regionSouth")
plot(political_preferences1, "trace", "regionWest")
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-trace3
#| fig-cap: "Trace plot of employ"
#| fig-subcap: ["Trace plot of employ Homemaker", "Trace plot of employ Part-time", "Trace plot of employ Permanently disabled", "Trace plot of employ Retired", "Trace plot of employ Student", "Trace plot of employ Temporarily laid off", "Trace plot of employ Unemployed", "Trace plot of employ Other"]
#| layout-ncol: 2

plot(political_preferences1, "trace", "employHomemaker")
plot(political_preferences1, "trace", "employPart-time")
plot(political_preferences1, "trace", "employPermanently disabled")
plot(political_preferences1, "trace", "employRetired")
plot(political_preferences1, "trace", "employStudent")
plot(political_preferences1, "trace", "employTemporarily laid off")
plot(political_preferences1, "trace", "employUnemployed")
plot(political_preferences1, "trace", "employOther")
```

@@

## Credibility Interval {#sec-credibility-interval}
90% Credibility intervals graph are shown in @fig-modelresults1 and @fig-modelresults2.
